---
title: "The effect of intercept changes on shift detection"
author: "Cecile Ane and Qing Yu(Sabrina)"
date: "October 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal of the study
In this report, we use `Rstudio` with version `r getRversion()` and `l1ou` package with version `r packageVersion("l1ou")`. Our goal is to explore the changes on the shift detection once the intercept has changed. From the `bs_shift` report, we can see the number of false-positives (the number of non-shift edges which are detected as shift edges) close to the root decreased when we set the intercept as 0. The reason is that if the intercept is not zero, the root edge is considered as one of the shifting edges. Now, the result shows that the existence of the intercept will influence the number of shift edges that are detected.


```{r,include=FALSE}
library('phylolm')
library('l1ou')
library(plotrix)
```

# Generating data for the first sample.
In this case, I do not include the last 10 tips on the tree.The `eModel` and `eModel2` are identical besides their intercepts. I try to compare if the shftedges detected by the two models are different or not. 

This is to check whether `lizard.tree` and `lizard.traits` correspond to the same sequence of species. 
```{r}
data(lizard.tree, lizard.traits)
all.equal(names(lizard.traits[,1]),lizard.tree$tip.label)
```
It is shown that the sequence of the tip labels are the same for `lizard.tree` and `lizard.traits`.

Next, we move on to check if the two models with different intercepts have different shift configurations.

The function `drop.tip` is used to exclude the last 10 tips on the tree. 
```{r, eval=FALSE}
data(lizard.tree, lizard.traits)
tr1=drop.tip(lizard.tree,91:100)
trait1=lizard.traits[,1][-(91:100)]
lizard <- adjust_data(tr1,trait1)
eModel <- estimate_shift_configuration(lizard$tree, lizard$Y)
eModel2 <- estimate_shift_configuration(lizard$tree, lizard$Y+400)
save(eModel, file="eModelint.RData")
save(eModel2, file="eModelint2.RData")
```


## Check if the shift edges for two models are the same or not for the first model
```{r}
load('eModelint.RData')
load('eModelint2.RData')
all.equal(sort(eModel$shift.configuration),sort(eModel2$shift.configuration))
```
The intercept for `eModel` is `r eModel$intercept` and the intercept for `eModel2` is `r eModel2$intercept`. The shift edges that has been detected from the two models are the same.


# Generating data for the second sample.
In this case, I do not include the 81st to 90th tips on the tree.The `eModel` and `eModel2` are identical besides their intercepts. I try to compare if the shftedges detected by the two models are different or not. 

 
```{r, eval=FALSE}
data(lizard.tree, lizard.traits)
tr2=drop.tip(lizard.tree,81:90)
trait2=lizard.traits[,1][-(81:90)]
all.equal(names(trait2),tr2$tip.label)
lizard <- adjust_data(tr2,trait2)
eModel <- estimate_shift_configuration(lizard$tree, lizard$Y)
eModel2 <- estimate_shift_configuration(lizard$tree, lizard$Y+400)
save(eModel, file="eModelint21.RData")
save(eModel2, file="eModelint22.RData")
```

## Check if the shift edges for two models are the same or not for the second model
The intercept for `eModel` is `r eModel$intercept` and the intercept for `eModel2` is `r eModel2$intercept`.
We also calculated the theta values for all shift edges to see how different they are.

```{r myplot,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
load('eModelint21.RData')
load('eModelint22.RData')
all.equal(sort(eModel$shift.configuration),sort(eModel2$shift.configuration))
shiftedge1=eModel$shift.configuration
shiftedge2=eModel2$shift.configuration
shiftedge1
shiftedge2
layout(matrix(1:2, 1, 2))
plot(eModel$tree,show.tip.label = F,main="Tree of eModel with shifts detected",font=8)
edgelabels(shiftedge1,shiftedge1,col="blue",frame="none")
plot(eModel2$tree,show.tip.label = F,main="Tree of eModel2 with shifts detected",font=8)
edgelabels(shiftedge2,shiftedge2,col="green",frame="none")

```

The shift edges that has been detected from the two models are not the same.There are less shift edges that have been detected for `eModel2` which has a higher intercept. 

# Generating data for the third sample.
In this case, I do not include the first 10 tips on the tree.The `eModel` and `eModel2` are identical besides their intercepts. I try to compare if the shftedges detected by the two models are different or not. 

```{r, eval=FALSE}
data(lizard.tree, lizard.traits)
tr3=drop.tip(lizard.tree,1:10)
trait3=lizard.traits[,1][11:100]
all.equal(names(trait3),tr3$tip.label)
lizard <- adjust_data(tr3,trait3)
eModel <- estimate_shift_configuration(lizard$tree, lizard$Y)
eModel2 <- estimate_shift_configuration(lizard$tree, lizard$Y+700)
save(eModel, file="eModelint31.RData")
save(eModel2, file="eModelint32.RData")
```

## Check if the shift edges for two models are the same or not for the third model
The intercept for `eModel` is `r eModel$intercept` and the intercept for `eModel2` is `r eModel2$intercept`.
```{r myplot2,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
load('eModelint31.RData')
load('eModelint32.RData')
all.equal(sort(eModel$shift.configuration),sort(eModel2$shift.configuration))
shiftedge3=eModel$shift.configuration # Those are the indices of shift edges
shiftedge4=eModel2$shift.configuration 
truetheta1 = eModel$intercept + l1ou:::convert_shifts2regions(eModel$tree, 
                                eModel$shift.configuration, eModel$shift.values)

truetheta2 = eModel2$intercept + l1ou:::convert_shifts2regions(eModel2$tree, 
                                eModel2$shift.configuration, eModel2$shift.values)

layout(matrix(1:2, 1, 2))
plot(eModel$tree,show.tip.label = F,main="Tree of eModel with shifts detected",font=8)
edgelabels(shiftedge3,shiftedge3,col="blue",frame="none")
plot(eModel2$tree,show.tip.label = F,main="Tree of eModel2 with shifts detected",font=8)
edgelabels(shiftedge4,shiftedge4,col="green",frame="none")
```
The shift edges that has been detected from the two models are not the same.There are less shift edges that have been detected for `eModel2` which has a higher intercept. 

Now, shift edge 14,55,74,164 has not been detected in `eModel2`. Let us check its theta value in `eModel` and `eModel2`compared to its neighbours` theta valueS.
```{r}
truetheta1[12:16] #shift edge 14
truetheta2[12:16] #shift edge 14
truetheta1[54:57] # shift edge 55
truetheta2[54:57] # shift edge 55
truetheta1[73:76] # shift edge 74
truetheta2[73:76] # shift edge 74
truetheta1[164:167] # shift edge 164
truetheta2[164:167] # shift edge 164
```
If we take a look at the truetheta values of shift edges and their neighbours, we can see a clear difference for the shift edge from its neighbours in `eModel`. However, the difference has disappeared when we look at `eModel2`. We should further investigate the reason for the disappearance of difference for shift edges with its neighbours.
