---
title: "Bootstrap shift support by different procedures"
author: "Cecile Ane and Qing (Sabrina) Yu"
date: "Oct 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal of the study

In this report, my goal is to find the procedure that gives "most accurate" support values for shifts.
I first estimate the shift configuration on a dataset based on one of 3 criteria: "AICc","BIC","pBIC" (first criterion).
Then, this model is used by the bootstrap procedure to simulate bootstrap data.
Each bootstrap data is analyzed with AICc, BIC or pBIC (second criterion).
There are 9 combinations of criteria (first + second). I compared the bootstrap support values from these nine procedures, on simulated data described below. I found that by using AIC + pBIC,
I got some of the 8 shifts with high bootstrap support and rare high support 
for wrong edges (edges where no shifts were simulated). 

 
AICc + pBIC gave high support at an acceptable rate (for 0.5 wrong edges on average)
Other procedures gave high support for less than 0.5 wrong edges on average,
but they gave low support for the true shifts.
So, it seems that AICc_pBIC is the "most accurate" procedure in a conservative way.  



# Simulation procedure

I used the lizard tree from the `phylolm` package and associated trait data on these lizard
species (just the first trait, which is the first PC axis from a PCA).
I analyzed this trait to estimate the shifts in trait evolution using the function
`estimate_shift_configuration` from `l1ou` package. The analysis takes a little bit of
time (about 2 minutes), so I saved the results in an R data file:

```{r eval=FALSE,include=TRUE}
setwd("~/R/traitevoOUshift/bootstrap")
library(l1ou)
data(lizard.tree, lizard.traits)
lizard <- adjust_data(lizard.tree, lizard.traits[,1])
eModel <- estimate_shift_configuration(lizard$tree, lizard$Y)
save(eModel,file="eModel.RData")
```

The estimated shift configuration can be loaded quickly now:

```{r, include=FALSE}
setwd("~/R/traitevoOUshift/bootstrap")
require('ape')
require('l1ou')
library(plotrix)
load('Y_res2016_09_02.RData')
load('boot_res2016_09_30.RData')
load('seed_res2016_09_02.RData')
load("eModel2016_09_02.RData")

```


```{r}
critlist = c("AICc","BIC","pBIC")
n_edges=length(eModel$tree$edge.length) # Total number of edges
n_edges
shiftedge=eModel$shift.configuration # Those are the indices of shift edges
shiftedge
otheredge=(1:n_edges)[-shiftedge] # other edges: those without a true shift
head(otheredge)
truealpha=eModel$alpha
# truealpha=truealpha/4
# y0=0
y0=eModel$intercept
truetheta   = l1ou:::convert_shifts2regions(eModel$tree, 
  eModel$shift.configuration, eModel$shift.values)+y0
nShifts   = length(eModel$shift.configuration)
n_tips=length(eModel$tree$tip.label)
critlist = c("AICc","BIC","pBIC")
sigma2=eModel$sigma2
```
The tree has `r n_tips` tips, and so `r n_edges` edges total.
There are `r nShifts` edges that were detected with a shift, numbered 
`r shiftedge`. On these edges, the changes in the optimal value were
(shift magnitude):
`r eModel$shift.values`.
The other parameters were estimated to be $\alpha=$ `r truealpha`, 
and $\sigma^2=$ `r sigma2`.

Now I used this model to simulate new data. Then, I analyzed each simulated data set with 
each of the 9 bootstrap procedures described in the introduction. I named this procedure
based on the first and second criteria (to analyze the original data and then to analyze
its bootstrap replicates):
`aic_aic`, `aic_bic`, `aic_pbic`, `bic_aic`, `bic_bic`, `bic_pbic`, `pbic_aic`, `pbic_bic`, and `pbic_pbic`.
The code to run the simulation is shown below. 
In particular, I choose to set the maximum allowed number of shifts to 20 instead of the 
default (half the number of tips) to make the simulation run in a reasonable time frame.

I divided the analysis of 100 simulated datasets into 10 sets of
10 datasets, and I used 10 different machines (each handled 1 set of 10 data sets).
This simulation took 18 hours for each machine. I used 4 cores from each machine, through
the options `multicore` and `nCores` of the bootstrapping function.

```{r, eval=FALSE, include=TRUE}
## simulation using bootstrap method
mnshifts = 20 # allowed maximum number of shifts. 
# The default is 50 = 100 tips/2 but too slow, 5 sim *5boot_itr took 40 hours.
mnshift_boot = 20 # check if it doesn't take too long
## simulation using bootstrap method
do.boot.sim = function(n_sim , nbootItr, seed, filename, numcores=1){
  srand(seed)
  #initiate all tables
  boot_pro=matrix(nrow = (9*n_sim), ncol=n_edges+2,data=NA) # two extra columns for criterion1 and 2
  irow=1
  Y_table=matrix(data=NA,nrow=n_sim,ncol=n_tips)
  
  for (i in 1:n_sim){
    Y  <- rTraitCont(eModel$tree, "OU", theta=truetheta, 
                     alpha=truealpha, 
                     sigma=sqrt(eModel$sigma2),root.value = y0)
    Y_table[i,]=Y
    #change different criterion when bootstrapping
    for (criterion1 in critlist){
      bootmodel=estimate_shift_configuration(eModel$tree,Y,criterion=criterion1, max.nShifts=mnshifts)
      for (criterion2 in critlist){
        bootmodel$l1ou.options$criterion = criterion2
        bootmodel$l1ou.options$max.nShifts = mnshift_boot
        bootB=l1ou_bootstrap_support(bootmodel, nItrs=nbootItr, multicore=T, nCores=numcores)
        boot_pro[irow,1]=criterion1
        boot_pro[irow,2]=criterion2
        boot_pro[irow,3:(n_edges+2)]=bootB
        irow=irow+1
      }
    }
  }
  save(seed,boot_pro,Y_table, file= filename)
  return (boot_pro)
}
# Call this function 10 times, each time to do 10 simulations with 100 iterations per simulation 
myresultsL <- do.boot.sim(100, 100, seed=1235, filename="test", numcores=4)
# seeds that were used for the 10 runs:
# seed        1235
# seed     127643435
# seed      12343445
# seed      12332375
# seed      12787875
# seed      21565385
# seed       7154895
# seed         95475
# seed     651789823
# seed         12345
```


# Results: summaries and visualizations

I first created 9 tables, one for each of the 9 procedures,
to contain the bootstrap value for each edge from each simulated dataset.
With 100 simulated data sets and 198 edges, that makes 9 tables each with
100 rows and 198 columns.

```{r}
boot_pro=as.data.frame(boot_rep)
for( j in 3:200){
  boot_pro[,j]=as.numeric(as.character(boot_pro[,j]))
}
# delete the first two columns, which contain the labels of different procedures, like AICc or BIC
aic_aic = boot_pro[ boot_pro[,1]=="AICc" & boot_pro[,2]=="AICc" ,-(1:2) ]
aic_bic = boot_pro[ boot_pro[,1]=="AICc" & boot_pro[,2]=="BIC" ,-(1:2)  ]
aic_pbic = boot_pro[ boot_pro[,1]=="AICc" & boot_pro[,2]=="pBIC" , -(1:2) ]
bic_aic = boot_pro[ boot_pro[,1]=="BIC" & boot_pro[,2]=="AICc" , -(1:2) ]
bic_bic = boot_pro[ boot_pro[,1]=="BIC" & boot_pro[,2]=="BIC" , -(1:2) ]
bic_pbic = boot_pro[ boot_pro[,1]=="BIC" & boot_pro[,2]=="pBIC" , -(1:2) ]
pbic_aic = boot_pro[ boot_pro[,1]=="pBIC" & boot_pro[,2]=="AICc" , -(1:2) ]
pbic_bic = boot_pro[ boot_pro[,1]=="pBIC" & boot_pro[,2]=="BIC" , -(1:2) ]
pbic_pbic = boot_pro[ boot_pro[,1]=="pBIC" & boot_pro[,2]=="pBIC" , -(1:2) ]

# creating the vector of tables
n=9
vectorOfTables <- vector(mode="list",length=n)
vectorOfTables[[1]]=aic_aic
vectorOfTables[[2]]=aic_bic
vectorOfTables[[3]]=aic_pbic
vectorOfTables[[4]]=bic_aic
vectorOfTables[[5]]=bic_bic
vectorOfTables[[6]]=bic_pbic
vectorOfTables[[7]]=pbic_aic
vectorOfTables[[8]]=pbic_bic
vectorOfTables[[9]]=pbic_pbic

plot_name=c("AICc_AICc","AICc_BIC","AICc_pBIC","BIC_AICc","BIC_BIC",
            "BIC_pBIC","pBIC_AICc","pBIC_BIC","pBIC_pBIC")
# show first 6 rows and 10 columns of aic_aic and aic_pbic tables
head(aic_aic[,1:10])
head(aic_pbic[,1:10])
```


I summarized the bootstrap support (BS) values by their mean, standard deviation, minimum and maximum
across the 100 simulations, for each edge separately.

#### Mean BS and other summaries, for shift edges
```{r}
means_sum=matrix(data=NA,nrow=9,ncol=length(shiftedge)) # mean of BS of shift edges.
sds_sum=matrix(data=NA,nrow=9,ncol=length(shiftedge)) # standard deviation of BS of shift edges.
mins_sum=matrix(data=NA,nrow=9,ncol=length(shiftedge)) # minimal number of BS of shift edges.
all_means=matrix(data=NA,nrow=9,ncol=n_edges) # mean of BS of all edges across 100 simulations.

for (i in 1:9){
  all_means[i,]=colMeans(vectorOfTables[[i]])
  means_sum[i,]=colMeans(vectorOfTables[[i]][,shiftedge])
  sds_sum[i,]=apply(vectorOfTables[[i]],2,sd)[shiftedge]
  mins_sum[i,]=apply(vectorOfTables[[i]],2,min)[shiftedge]
}
round(means_sum,3) # Mean BS: row = procedure (like aic_aic), column = edge with shift
round(sds_sum,3)  # SD of BS: row = procedure (like aic_aic), column = edge with shift
```

Now we can visualize these BS means, one for each edge:

```{r}
matplot(means_sum, type = c("o"),pch=1,col = 1:8,ylab="mean BS for shift edges",axes=F)
axis(side=1,at=1:9,labels = plot_name,cex.axis=0.7)
axis(side=1,at=2,labels = plot_name[2],cex.axis=0.7)
axis(side=1,at=3,labels = plot_name[3],cex.axis=0.7)
axis(side=1,at=4,labels  = plot_name[4],cex.axis=0.7)
axis(side=1,at=7,labels = plot_name[7],cex.axis=0.7)
axis(side=1,at=9,labels = plot_name[9],cex.axis=0.7)
axis(side=2,ylim=c(0:0.6))
mtext("mean bootstrap support on edges that truly had a shift. 1 line = 1 edge",
      line=1,side=3,las=0)
```

We see that AICc + AICc gives the highest support for shifts on these edges. 
We also see that all bootstrap support values tend to be low (given that shifts were simulated on these
edges).
This might be because the magnitudes of the shifts were small.
But there are systematic differences between edges: two of them (in yellow and pink) tend 
to be supported much more than the others. Another two (green and gray) are almost never
supported (with mean BS<0.1 across all procedures). 

#### Mean BS and other summaries, for non-shift edges

```{r}
meano_sum=matrix(data=NA,nrow=9,ncol=n_edges-length(shiftedge))
sdo_sum=matrix(data=NA,nrow=9,ncol=n_edges-length(shiftedge))
maxo_sum=matrix(data=NA,nrow=9,ncol=n_edges-length(shiftedge))
# will contain the mean, standard deviation, max bootstrap support
# of non-shift edges across 100 simulations.

# now filling in the values
for (i in 1:9){
  meano_sum[i,]=colMeans(vectorOfTables[[i]][,otheredge])
  sdo_sum[i,]=apply(vectorOfTables[[i]],2,sd)[otheredge]
  maxo_sum[i,]=apply(vectorOfTables[[i]],2,max)[otheredge]
}

round(meano_sum[,1:10],3) # Mean BS: row = procedure (like aic_aic), column = edge with no shift
round(sdo_sum[,1:10],3) # SD BS: row = procedure (like aic_aic), column = edge with no shift
```

Now we can again visualize these BS means, one for each edge:

```{r}
#Plotting the means,sds
layout(matrix(1, 1))
matplot(meano_sum, type = c("o"),pch=1,ylab="mean BS of non-shift edges",axes=F)
axis(side=1,at=1:9,labels = plot_name,cex.axis=0.7)
axis(side=1,at=2,labels = plot_name[2],cex.axis=0.7)
axis(side=1,at=3,labels = plot_name[3],cex.axis=0.7)
axis(side=1,at=4,labels = plot_name[4],cex.axis=0.7)
axis(side=1,at=7,labels = plot_name[7],cex.axis=0.7)
axis(side=1,at=9,labels = plot_name[9],cex.axis=0.7)
axis(side=2,ylim=c(0:0.6))
mtext("mean bootstrap support on edges that do not have a shift. 1 line = 1 edge",
      line=3,side=3,las=0)
```
We see quite a high bootstrap support overall when the second criterion is AICc.
We also see that 2 edges stand out (in green and pink) as having very high bootstrap support
(up to 40%), even though there were no shifts simulated on these edges.
The two edges are shown below:

```{r}
which(meano_sum[1,]>.2) #  189th in "otheredge" = 197 in table with all edges.
```
They are both edges close to the tips (based on the plot below).
These two edges have bootstrap support `r meano_sum[1,30]` and `r meano_sum[1,189]` with AICc+AICc
one average. We can look at the distribution (not just the mean) of their BS across the 100 simulations:

```{r}
layout(matrix(1:4, 2, 2, byrow=TRUE))
#hist(aic_aic[,otheredge[30]],breaks=20,xlab="bootstrap support",
 #    main="edge 31 (no shift), AIC+AIC",col="tan")# almost flat between 0 and 1
#hist(aic_pbic[,otheredge[30]],breaks=20,xlab="bootstrap support",
#     main="edge 31, AIC+pBIC ", col="green", xlim=0:1)# unimodal near 0
hist(aic_aic[,otheredge[189]],breaks=20,xlab="bootstrap support",
     main="edge 197 (no shift), AIC+AIC",col="tan")# almost flat between 0 and 1
hist(aic_pbic[,otheredge[189]],breaks=20,xlab="bootstrap support",
     main="edge 197, AIC+pBIC",col="green", xlim=0:1)# unimodal near 0
```
These histrograms are uni-modal. In the plots of AICc + AICc,
BS is almost uniformly distributed from 0 to 1. With AIC + pBIC, the BS values are much lower, most of
them near BS=0, but still unimodal. 

### Distribution of BS on shift edges to find a threshold for "high" versus "low" support

Next, we can look at the distribution of bootstrap support on edges with
a true shift, to see what threshold would be most appropriate to call "high" support,
given that our shifts were of small magnitude and were difficult to detect
during the bootstrap procedure. We are going to look at the 2 shifts that were easiest to detect.

```{r}
layout(matrix(1:4, 2, 2, byrow=TRUE))
hist(aic_aic[,shiftedge[4]],breaks=20,xlab="bootstrap support",
     main="edge 74 (with shift), AIC+AIC",col="tan")
hist(aic_pbic[,shiftedge[4]],breaks=20,xlab="bootstrap support", xlim=0:1,
     main="edge 74, AIC+pBIC ", col="green")
hist(aic_aic[,shiftedge[5]],breaks=20,xlab="bootstrap support",
     main="edge 14 (with shift), AIC+AIC",col="tan")
hist(aic_pbic[,shiftedge[5]],breaks=20,xlab="bootstrap support",
     main="BS of edge 14, AIC+pBIC",col="green")
```

These plots are all bimodal: with a big chunk at/near 0, and another between 0.4 to 1.
These plots make us think that a BS of 0.4 could be a useful cut-off point to differentiates
high versus low bootstrap support. Another reason for this is that for edges with no shifts 
(previous section), few simulated data had BS below 0.4.




# Display of edge BS means and frequency of high support

Here we display results for each edge, summarized across all simulated data. 
Below, the plot on the left shows the mean bootstrap support for each edge, if it is 0.05 or higher.
The plot on the right shows, for each edge, the proportion of replicates who had a BS of 40% of higher
for that edge (this proportion is not shown if it is lower than 0.01).

These plots are shown for 2 procedures only: the most liberal AICc+AICc and our recommended
procedure: AICc+pBIC.  

```{r myplot,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
plotShiftSupport = function(k){
  layout(matrix(1:2, 1, 2))
  bs_high=colSums(vectorOfTables[[k]]>=0.4)
  pro_high=bs_high/100
  ind_high=which(pro_high>0.01)
  ind_means=which(all_means[k,]>=0.05)
  ins_ps=intersect(shiftedge,ind_high)
  ins_po=intersect(otheredge,ind_high)
  ins_ms=intersect(shiftedge,ind_means)
  ins_mo=intersect(otheredge,ind_means)
  list =c(1,2,4,7) # those methods have indices from ind_means that are not shift edges.
  plot(eModel$tree,show.tip.label = F,main=plot_name[k],sub=
"Blue: edges with a true shift. 
Pink: edges with no shift.",font=8)
  edgelabels(round(all_means[k,ins_ms],2),ins_ms,col="blue",frame="none")
  if (k %in% list){
  edgelabels(round(all_means[k,ins_mo],2),ins_mo,col="pink",frame="none")
  }
  mtext(text ="Mean bootstrap support across simulations.
Not shown if < 0.05.",font=6,line=-1)
  plot(eModel$tree,show.tip.label = F,main=plot_name[k],sub=
"Red: edges with a true shift. 
Green: edges with no shift.",font=8)
  edgelabels(round(pro_high[ins_ps],2),ins_ps,col="red",frame="none")
  edgelabels(round(pro_high[ins_po],2),ins_po,col="green",frame="none")
  mtext(text ="Proportion of replicates with BS >=0.4.
Not shown if <= 0.01.", line=-1)
}
plotShiftSupport(1)
```

For AICc+AICc, 2 shift edges have BS>=0.4 in a reasonable proportion of simulated data: 57% and 62%
(still, this is not very high).
Another 4 shift edges have a BS>=0.4 between 21% and 32% of the replicates.
The remaining 2 edges with a shift have "high" support in 12% or less of the replicates.
Many non-shift edges have "high" BS in 6% or more of the simulated data.
The 2 weird non-shift edges have "high" support in 40% and 44% of the replicates,
which is very high compared to the support found for shift edges.

```{r myplot4, echo=TRUE, fig.height=12, fig.width=8, dpi=300}
plotShiftSupport(3)
```

For AICc+pBIC, only one non-shift edge has BS>=0.4 in 6% or more of the simulated data.
The 2 best support shift edges have high BS in 25% and 62% of the simulated data.
All remaining 6 shift edges have BS above 0.4 in 8% or less of the replicates.

For the figures for all the 9 procedures, please see file `prop_plots.pdf`.

```{r myplot3, eval=FALSE, include=TRUE}
pdf("prop_plots.pdf", width=8.5, height=11)
layout(matrix(1:2, 1, 2))
for (k in 1:9){ # kth row = kth procedure. k=1=the first row=aic_aic
  plotShiftSupport(k)
}
dev.off()
```
The results for the  procedures not shown above are summarized below:

For AICc+BIC, the 2 shifts that were easiest to detect both had the mean bootstrap support of 0.13.

When the first criterion is BIC, many non-shift edges have high support in 5% or more of the
simulation replicates. One of the shift edges has high support in <= 1% of the replicates.
The 2 most highly supported shifts have mean BS of 0.23 and 0.19 when the second criterion is AICc.

When pBIC was the first criterion, non-shift edges tended to have low support,
but 6 true shifts had high BS in 2% of less of simulated data.


# Overall number of false and true positives

We now look at all edges with high support for each given replicate.
We define a false positive as a non-shift edge that had the bootstrap support of 0.4 or higher.
A true positive is a shift edge with the bootstrap support of 0.4 or higher.
For each simulated data, we counted the number of false positives and the number
of true positives, and then we calculated the mean and standard deviation of these
false and true positives across 100 replicates.

```{r}
falist=matrix(data=NA,nrow=9,ncol=100) # number of non-shift edges with high BS: row=procedure, col=simulated data
trulist=matrix(data=NA,nrow=9,ncol=100) # number of shift edges  with high BS
for (k in 1:9){
  for (i in 1:100){
    falist[k,i]=sum(vectorOfTables[[k]][i,otheredge]>=0.4)
    trulist[k,i]=sum(vectorOfTables[[k]][i,shiftedge]>=0.4)
  }
}

# mean and sd of true and false-positive number for each procedure
mean_fal=rowMeans(falist)
sd_fal=apply(falist,1,sd)
mean_tru=rowMeans(trulist)
sd_tru=apply(trulist,1,sd)

summ_table=data.frame(meanFP=mean_fal, sdFP=sd_fal, meanTP=mean_tru, sdTP=sd_tru)
summ_table$crit1=c("AICc","AICc","AICc","BIC","BIC","BIC","pBIC","pBIC","pBIC")
summ_table$crit2=c("AICc","BIC","pBIC","AICc","BIC","pBIC","AICc","BIC","pBIC")
summ_table

xpos = c(1,2,3)
j=c(1,4,7) # procedures with aic as the second criterion
layout(matrix(1, 1))
interaction.plot(summ_table[,5],summ_table[,6],mean_fal,col=c("blue","gray","red"),
                 main="Mean number of false positives",
                 xlab="first criterion: on original data",
                 ylab="mean # of false positives",
                 trace.label="second criterion:\non bootstrap data",ylim=c(0,6))

# Add confidence interval on each point
for (m in critlist) {
  if (m=="AICc"){ 
    add=F
    meanplot=plotCI(xpos+0.03*(1-0.5*add),mean_fal[j],2*sd_fal[j]/10,
                    main ="", axes=F, add=T,pch=15,sfrac=0,col="blue")
  }
  if (m == "BIC") {
    add=T
    meanplot=plotCI(xpos+0.05*(1-add),mean_fal[j+1],2*sd_fal[j+1]/10,
                    main ="",ylab="",add=add,pch=16,sfrac=0,col="gray")
  } 
  if (m=="pBIC"){ add=T
  meanplot=plotCI(xpos+0.05*(1-1.5*add),mean_fal[j+2],2*sd_fal[j+2]/10,
                  main ="",axes=F, add=T,pch=17,sfrac=0, col="red")
  }
}

interaction.plot(summ_table[,5],summ_table[,6],mean_tru,col=c("blue","gray","red"),
                 main="Mean number of true positives",
                 xlab="first criterion: on original data",
                 ylab="mean # of true positives",
                 trace.label="second criterion:\non bootstrap data",ylim=c(0,6))
# Add confidence interval on each point
for (m in critlist) {
  if (m=="AICc"){ 
    add=F
    meanplot=plotCI(xpos+0.03*(1-0.5*add),mean_tru[j],2*sd_tru[j]/10,
                    main ="", axes=F, add=T,pch=15,col="blue",sfrac=0)
  }
  if (m == "BIC") {
    add=T
    meanplot=plotCI(xpos+0.05*(1-add),mean_tru[j+1],2*sd_tru[j+1]/10,
                    main ="",ylab="",add=add,pch=16,col="gray",sfrac=0)
  } 
  if (m=="pBIC"){ add=T
  meanplot=plotCI(xpos+0.05*(1-1.5*add),mean_tru[j+2],2*sd_tru[j+2]/10,
                  main ="",axes=F, add=T,pch=17,sfrac=0,col="red")
  }
}

```

When AICc is the first criterion, then the second criterion can change the mean number of false positives
and true positives significantly: with AICc as the second criterion again, the mean number of false positive 
and of true positives is larger than with BIC or pBIC.
When BIC or pBIC is the first criterion, the second criterion has little impact on the number of false/true positives. There are fewer true positives with pBIC than with BIC (as first criterion), for a similar number of false positives. 

The distribution of false and true positives is unimodal in all cases. Only 2 procedures are shown below:
the most liberal AICc+AICc and our recommended procedure AICc+pBIC:

```{r}
layout(matrix(1:2, 1, 2))
# create histgram of false-positive BS for each procedure
hist(falist[1,],main=plot_name[1],breaks=40,col=1,sub="Number of false positives",xlab="")
hist(falist[3,],main=plot_name[3],breaks=40,col=3,sub="Number of false positives",xlab="")

# create histgram of true-positive BS for each procedure
hist(trulist[1,],main=plot_name[1],breaks=40,col=2,sub="Number of true positives",xlab="")
hist(trulist[3,],main=plot_name[3],breaks=40,col=4,sub="Number of true positives",xlab="")
```


Most of the time, 
AICc + AICc lead to 2-9 false positives and 1-4 true positives, and
AICc + pBIC lead to 0-2 false positives and 0-2 true positives.

See files `falist_plots.pdf` and  `trulist_plots.pdf` for the false and true positive histograms with other procedures.

```{r myplot2, eval=FALSE, include=TRUE}
# create histgram of false-positive BS for each procedure
pdf("falist_plots.pdf", height=10.5, width=8.5)
par(mfrow=c(2,2))
for (k in 1:9){
  hist(falist[k,],main=plot_name[k],breaks=40,col=k,sub="Number of false positives",xlab="")
}
dev.off()

# create histgram of true-positive BS for each procedure
pdf("trulist_plots.pdf", height=10.5, width=8.5)
par(mfrow=c(2,2))
for (k in 1:9){
  hist(trulist[k,],main=plot_name[k],breaks=40,col=k,sub="Number of true positives",xlab="")
}
dev.off()
```



