---
title: "Estimate contrast with known alpha"
author: "Cecile Ane and Qing (Sabrina) Yu"
date: "Feb 22,2017"
output:
  pdf_document: default
  toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal of the study

In this report, our goal is to check the distribution of phylogenetic contrasts
with known alpha, known shift configuration and unknown shift values that are used in the bootstrap procedure 
in the `l1ou` package. In the last report, we used the true shift edge, true shift values, true alpha.

We used code from the bootstrap procedure to simulate contrast values.
Then we calculated the mean and standard deviation of the contrast values at each node.
The conclusions above held for all nodes, whether nodes had a shift on a child edge or not. 


#  Estimate shift configuration from original lizard data (change)

We used the lizard tree from the `phylolm` package and associated trait data on these lizard
species (just the first trait, which is the first PC axis from a PCA).
To simulate data according to a model that is biologically plausible,
we estimated the shift configuration, shift values and covariance parameters
from original lizard data.

```{r,include=FALSE,eval=FALSE}
eModel <- estimate_shift_configuration(lizard$tree, lizard$Y)
save(eModel,  file = "eModel_2_27.RData")
```

This trait was analyzed to estimate the shifts in trait evolution using the function
`fit_OU` from the `l1ou` package. 
```{r,i, include=FALSE}
library(l1ou)
library(phylolm)
```

```{r}
sessionInfo()$otherPkgs$l1ou$Version  #version 1.29
```


```{r}

load("eModel_2_27.RData") # eModel.tree same as lizard.tree
data(lizard.tree, lizard.traits)
lizard <- adjust_data(lizard.tree, lizard.traits[,1])
n_tips=length(lizard$tree$tip.label)  # 100: total number of tips
truealpha=eModel$alpha  # 0.606897967627925
y0=eModel$intercept
truetheta = y0 + l1ou:::convert_shifts2regions(eModel$tree, 
                                eModel$shift.configuration, eModel$shift.values)
table(truetheta)
# -5.25186223573345 -4.56654712450638 -4.04766913950491 -3.45016309674712 -3.23947780027265 -2.72293353502748 
#                1                 3                 1                 1                 1                 7 
#-2.36362306302032  -1.6605679645554 0.248809571469948 
#                1                25               158 
nShifts   = length(eModel$shift.configuration) # c(55,98,118,74,14,77,32,164)

sigma2=eModel$sigma2   # 0.0625186613743141
shift_config=eModel$shift.configuration
shiftnode=eModel$tree$edge[shift_config,1]# internal nodes with shift

```


Below, we use this model to simulate new data using the `rTraitCont` function from the `phylolm` package.
`RE` is the result of the function `sqrt_OU_covariance`, which calculates the square-root of the 
phylogenetic covariance matrix with a recursive algorithm, which traverses the tree once.
`covInverseSqrt` is the inverse square root of the phylogenetic covariance matrix, and `covSqrt` is the square-root of the phylogenetic covariance matrix. 
Finally, `contrast` contains the contrasts at all nodes. 
These matrices and contrasts were obtained using the true value of $\alpha$,
the same value used to simulate the data. This is an ideal situation when
$\alpha$ is known without error.

# simulation procedure
We use the same data for all scenarios:
`set.seed` is used here to make it possible to re-simulate the exact same data,
then these data are saved in a file (`Y_table`).

```{r,eval=F}
n_sim=100000
Y_table=matrix(nrow=n_sim, ncol=n_tips, data=NA)
set.seed(1293)
for (i in 1:n_sim) {
  Y  <- rTraitCont(eModel$tree, "OU", theta=truetheta, 
                 alpha=truealpha, 
                 sigma=sqrt(eModel$sigma2), root.value=y0)
  Y_table[i,]=Y
}
colnames(Y_table)=eModel$tree$tip.label
save(Y_table, file="Y_table_2_27.RData")
```

```{r}
load("Y_table_2_27.RData")
len=dim(Y_table)[1]
n_sim=100000
```

Just to check our simulated values, we define
our own function "ss" to calculate the variance of assuming a known population mean of 0.
The name "ss" stands for "sum of squares", although this sum of squares is rescaled by the sample size.
	
```{r}
ss=function(x){
	  sum(x^2)/length(x)
	}
```
	
Now to check the mean and variance of the simulated Y values:

```{r, eval=F}
round(colMeans(Y_table-rep(eModel$mu[,1],each=n_sim),5)) # all 0: good
convar = apply(Y_table-rep(eModel$mu[,1],each=n_sim), 2, ss)
round(convar,3) # all about the same
mean(convar) # 0.03622244: same as expected variance below: good
sigma2/(2*truealpha) * (1-exp(-2*truealpha*1)) # 0.03620576
```


# Contrast calculation procedure

We consider 4 scenarios for the calculation of contrasts, depending
on what parameters are known (yes) or unknown (no) for the calculation
of the inverse square-root of the covariance matrix.

parameter | 1: full model known | 2: alpha known | 3: shift config known | 4: tree known
----------|-----|-----|-----|-----
$\beta$<sub>0</sub>, $\beta$<sub>1</sub>, ...| yes | no | no | no
$\alpha$ | yes | yes | no | no
shift config | yes | yes | yes | no
For the tree topology, we assume it known always.
Refer to "filename" for the results of simulation 1, in which we know everything,
and the contrasts behave as expected.

## scenario 1: all parameters known and set to their true values

use Y table, call contrasts table "contrast_table1"
```{r,eval=F}
REf = sqrt_OU_covariance(lizard.tree, alpha=truealpha, 
                        root.model = "OUfixedRoot", # eModel$l1ou.options$root.model,
                        check.order=F, check.ultrametric=F) # needed in l1ou v1.29

covInverseSqrt  = t(REf$sqrtInvSigma)
```

To check the correctness of this square-root covariance matrix calculation:
```{r, eval=F}
treeheight = branching.times(eModel$tree)[1] # 1.00000000002
Vbm = vcv(eModel$tree) # t_ij values. Tree height is 1 here, do d_ij = 2*(1-t_ij)
Vou = (1-exp(-2*truealpha*Vbm)) * exp(-truealpha*2*(treeheight-Vbm)) / (2*truealpha)
tmp = covInverseSqrt %*%  Vou %*% t(covInverseSqrt)
diag(tmp) # if normalize.tree.height=T was ommitted above, and l1ou v1.29:
# all 1.742289: ratio found later in convar/sigma2
all.equal(tmp, diag(n_tips), scale=1) # TRUE

```

Now that we can trust the inverse square-root of the covariance matrix,
158	we use it to calculate contrasts:
```{r,eval=F}
contrast_table=matrix(nrow=n_sim, ncol=n_tips, data=NA)
for (i in 1:len) {
  contrast_table[i,]=covInverseSqrt%*%(Y_table[i,] - eModel$mu)
}
save(contrast_table, file="contrast_table_2_27.RData")
```



## scenario 2: beta unknown

Here, we assume that we know alpha, we know the shift configuration, but beta is unknown.
171	It took around 3 hours.

```{r,eval=F,include=T}
RE = sqrt_OU_covariance(lizard$tree, alpha=truealpha, # assumed known here
                        root.model =  eModel$l1ou.options$root.model,
                        check.order=F, check.ultrametric=F)
covInverseSqrt  <- t(RE$sqrtInvSigma)

contrast_table2=matrix(nrow=len, ncol=n_tips, data=NA)
sigma_table2=rep(NA,len)
mu_table2=matrix(nrow=len, ncol=n_tips, data=NA)
shift_values2=matrix(nrow=len, ncol=length(shift_config), data=NA)
for (i in 1:len) {
  model= fit_OU(lizard$tree, Y_table[i,], shift_config,
                alpha.upper=truealpha, alpha.lower=truealpha, alpha.starting.value=truealpha)
  # the line above is to fix alpha to the known value
 sigma_table2[i] = model$sigma2
 shift_values2[i,]=model$shift.values
 mu_table2[i,]=model$mu
  contrast_table2[i,]  <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(sigma_table2,file="knownalpha_sigma2.RData")
 save(contrast_table2,file="knownalpha_contrast.RData")
 save(mu_table2,file="knownalpha_mu.RData")
 save(shift_values2,file="knownalpha_shiftvalues.RData")
```

## scenario 3: alpha and beta unknown

Here we assume that the shift configuration is known, but not alpha and not beta.
The correlation matrix among tips is unknown, then, so its inverse square-root
needs to be re-calculated (based on the estimated alpha) for each simulated vector.
```{r,eval=F,include=T}
contrast_table3 = matrix(nrow=len, ncol=n_tips, data=NA)
alpha_table3  = rep(NA,len)
sigma_table3 = rep(NA,len)
mu_table3 = matrix(nrow=len, ncol=n_tips, data=NA)
shift_values3 = matrix(nrow=len, ncol=length(shift_config), data=NA)
# also save sigma2 estimated values, and 8 shift values? and/or 100 mu values?
for (i in 1:len) {
  model= fit_OU(lizard$tree, Y_table[i,], shift_config)
  alpha_table3[i] = model$alpha
  sigma_table3[i] = model$sigma2
  shift_values3[i,]=model$shift.values
  mu_table3[i,]=model$mu
  RE = sqrt_OU_covariance(lizard$tree, alpha=model$alpha, # alpha estimated here
                           root.model =  eModel$l1ou.options$root.model,
                          check.order=F, check.ultrametric=F)
  covInverseSqrt  <- t(RE$sqrtInvSigma)
  contrast_table3[i,] <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(contrast_table3,file="unknownalpha_contrast.RData")
save(alpha_table3, file="unknownalpha_alpha.RData")
save(sigma_table3,file="unknownalpha_sigma2.RData")
save(mu_table3,file="unknownalpha_mu.RData")
save(shift_values3,file="unknownalpha_shiftvalues.RData")
```

## scenario 4.1: unknown shift configuration

Also: unknown alpha and unknown beta.
The estimation of the shift configuration takes a long time, so the
number of simulations had to be reduced.
In a preliminary run, the contrasts for 1262 simulations were calculated in 24 hours.
So for the final run, we plan to complete a total of 2000 simulations ("n_sim" below).
criterion: pBIC, which is the default of function `estimate_shift_configuration`. 

```{r,eval=F,include=T}
n_sim=800 # lower because estimation is a lot slower when we have to search for the config
# use first n_sim rows of Ytable only
Y_table=Y_table[1:n_sim,]
contrast_table4=matrix(nrow=n_sim, ncol=n_tips, data=NA)
vectorOfShift <- vector(mode = "list", length = n_sim)
alpha_table4  = rep(NA,n_sim)
sigma_table4 = rep(NA,n_sim)
mu_table4 = matrix(nrow=n_sim, ncol=n_tips, data=NA)
shift_values4 = vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  model= estimate_shift_configuration(lizard$tree, Y_table[i,])
  vectorOfShift[[i]] =model$shift.configuration
  alpha_table4[i] = model$alpha
  sigma_table4[i] = model$sigma2
  shift_values4[[i]]=model$shift.values
  mu_table4[i,]=model$mu
  RE = sqrt_OU_covariance(lizard$tree ,alpha=model$alpha,
                          root.model = "OUfixedRoot",
                          check.order=FALSE, check.ultrametric=FALSE)
  covInverseSqrt  <- t(RE$sqrtInvSigma)
  contrast_table4[i,] <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(contrast_table4,file="unknownconfig_contrast_4.1.RData")
save(alpha_table4, file="unknownconfig_alpha_4.1.RData")
save(sigma_table4,file="unknownconfig_sigma2_4.1.RData")
save(mu_table4,file="unknownconfig_mu_4.1.RData")
save(shift_values4,file="unknownconfig_shiftvalues_4.1.RData")
save(vectorOfShift,file="unknownconfig_shifts_4.1.RData")
```

## scenario 4.2: unknown shift configuration, unknown alpha and unknown beta

criterion: AICc

```{r,eval=F,include=T}
n_sim=800 # lower because estimation is a lot slower when we have to search for the config
# use first n_sim rows of Ytable only
Y_table=Y_table[1:n_sim,]
contrast_table5=matrix(nrow=n_sim, ncol=n_tips, data=NA)
vectorOfShift2 <- vector(mode = "list", length = n_sim)
alpha_table5  = rep(NA,n_sim)
sigma_table5 = rep(NA,n_sim)
mu_table5 = matrix(nrow=n_sim, ncol=n_tips, data=NA)
shift_values5 = vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  model= estimate_shift_configuration(lizard$tree, Y_table[i,],criterion="AICc" )
  vectorOfShift2[[i]] =model$shift.configuration
  alpha_table5[i] = model$alpha
  sigma_table5[i] = model$sigma2
  shift_values5[[i]]=model$shift.values
  mu_table5[i,]=model$mu
  RE = sqrt_OU_covariance(lizard$tree ,alpha=model$alpha,
                          root.model = "OUfixedRoot",
                          check.order=FALSE, check.ultrametric=FALSE)
  covInverseSqrt  <- t(RE$sqrtInvSigma)
  contrast_table5[i,] <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(contrast_table5,file="unknownconfig_contrast_4.2.RData")
save(alpha_table5, file="unknownconfig_alpha_4.2.RData")
save(sigma_table5,file="unknownconfig_sigma2_4.2.RData")
save(mu_table5,file="unknownconfig_mu_4.2.RData")
save(shift_values5,file="unknownconfig_shiftvalues_4.2.RData")
save(vectorOfShift2,file="unknownconfig_shifts_4.2.RData")
```



# Results: visualizations of contrast distributions

```{r}
#nodes that corresponding to small variance contrasts
intnode=lizard.tree$edge[,1]
all.equal(intnode, rep(199:101, each=2))
intnode=unique(intnode) ##remove repetitive numbers
```
The sequence of internal node is from 199:101, which corresponding to contrasts from 1: 99. The 100th contrast is the contrast of the root node.
Here is a simple proof:
Now to check the mean and variance of the simulated Y values:
```{r}
# Now, we want to know which contrast does node 197 corresponding to. After we add 5 to all traits values in one sister branch and substract 5 from # all trait values in the other sister`s branch. The contrast of node 197 will change the most. Others may have slight changes or remain the same,
truealpha=eModel$alpha
REf = sqrt_OU_covariance(lizard.tree, alpha=truealpha, 
                        root.model =  eModel$l1ou.options$root.model,
                        check.order=F, check.ultrametric=F)

Dtf  = t(REf$sqrtInvSigma)
contrast_original=Dtf%*%(eModel$Y - eModel$mu)
eModel$Y[99]=eModel$Y[99]+5
eModel$Y[98]=eModel$Y[98]-5
eModel$Y[97]=eModel$Y[97]-5
eModel$Y[96]=eModel$Y[96]-5
contrast_change=Dtf%*%(eModel$Y - eModel$mu)
options(scipen=999)
cbind(contrast_original,contrast_change)
 
```

Here, we can see the third contrast changes the most significantly, so the node 197 corresponding to contrast 3. 

### Results from Scenario 1: all parameters known and set to their true values
```{r}
load("allknown_contrast.RData")
head(colMeans(contrast_table1)) 
max(abs(colMeans(contrast_table1)))# 0.0018224
round(colMeans(contrast_table1),5)
convar=apply(contrast_table1,2,ss)
round(convar,3)
mean(convar)# 0.06249321
sigma2# 0.06251866
0.1089751/0.06251866 # 1.743081. l1ou v1.29 without normalize.tree.height: mean(convar) was 0.1089751
```
Mean of contrasts are close to zero and the variance of contrasts is close to `sigma2`.


### Results from Scenario 2: ## scenario 2: beta unknown

but known alpha and known shift configuration
```{r myplot,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
load("knownalpha_contrast.RData")
load("knownalpha_sigma2.RData")
load("knownalpha_shiftvalues.RData")
load("knownalpha_mu.RData")
# contrasts mean
contrast_mean2=colMeans(contrast_table2)
head(contrast_mean2) 
max(abs(contrast_mean2))
hist(contrast_mean2)
# contrasts variance
convar2=apply(contrast_table2,2,ss)
contrast2=colMeans(contrast_table2)

#Try to find the contrast with variance relatively small, median and large. 
ind_small2=which(convar2<0.04)
ind_median2=which(convar2<0.06&convar2>0.04)
median_node2=intnode[ind_median2]
small_node2=intnode[ind_small2]
hist(convar2,main="variance of contrasts in Scenario 2",xlab="contrast variance",ylab="")
plot(eModel$tree,show.tip.label = F,main="",font=8)
nodelabels(round(convar2[ind_small2],3),small_node2,col="blue",frame = "none",cex=0.6)
edgelabels(edge=shift_config,col="green",frame = "none",pch=17)
nodelabels(round(convar2[ind_median2],3),median_node2,col="purple",frame = "none",cex=0.6)
mtext(

"blue: contrast variance < 0.04, purple:
      contrast variance > 0.04 and < 0.06,
      green dots: shifts",
      line=-21,side=3,las=0,cex=1)
plot(convar2~dist2shift, xaxt="n",xlab="distance to shift", ylab="contrast variance")
axis(1, at=0:4, labels=as.character(c(0:3,Inf)))
round(convar2,3)
mean(convar2)
mean(convar2[-ind_small2])
sigma2
```
Some contrasts with extremely small variance draw our attention. When we make a plot, we can see those contrasts with small variances appear where the shift exists. Those shifts make the variances of contrasts underestimated. And the plot below can clearly show as nodes become further away from shifts, their variances of contrast increase. The nodes closer to the root have higher variance of contrasts. And it is clear that the last contrast has variance equal to 0 since we have a fixed root.  
```{r}
###making plots for contrasts
dist2shift=rep(Inf,199)
dist2shift[shiftnode]=0
shiftparent=shiftnode
for(k in 1:10){
  old=c()
  for(node in shiftparent){
    ind=which(node==eModel$tree$edge[,2])
    old=c(old,ind)
  }
  shiftparent=eModel$tree$edge[old,1]
  for(node in shiftparent){
    if(dist2shift[node]>k) dist2shift[node]=k
  }
}

dist2shift=dist2shift[-c(1:99)]
dist2shift=rev(dist2shift)
dist2shift[100]=0 ##The last contrast is the root edge
dist2shift[dist2shift==Inf] = 4
plot(convar2~dist2shift, xaxt="n",xlab="distance to shift", ylab="contrast variance")
axis(1, at=0:4, labels=as.character(c(0:3,Inf)))
```

###Results from Scenario 3:  alpha and beta unknown
but known shift configuration
```{r}
load("unknownalpha_contrast.RData")
load("unknownalpha_sigma2.RData")
load("unknownalpha_shiftvalues.RData")
load("unknownalpha_mu.RData")
load("unknownalpha_alpha.RData")
# calculate contrasts mean
contrast_mean3=colMeans(contrast_table3)
head(contrast_mean3) 
max(abs(contrast_mean3))
hist(contrast_mean3)
# Calculate contrasts variance
convar3=apply(contrast_table3,2,ss)
round(convar3,3)
hist(convar3)
###Exactly the same as last one
intnode[convar3<0.06]
mean(convar3)
sigma2
#Try to find the contrast with variance relatively small, median and large. 
ind_small3=which(convar3<0.04)
ind_median3=which(convar3<0.06&convar3>=0.04)
ind_large3=which(convar3>=0.08)
median_node3=intnode[ind_median3]
small_node3=intnode[ind_small3]
large_node3=intnode[ind_large3]
large_node3
mean(convar3[-c(ind_small3,ind_large3)])
```

```{r myplot2,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
plot(eModel$tree,show.tip.label = F,main="small,median,large contrast nodes",font=8)
nodelabels(round(convar3[ind_small3],3),small_node3,col="blue",frame = "none")
edgelabels(shift_config,shift_config,col="green",frame = "none")
nodelabels(round(convar3[ind_median3],3),median_node3,col="purple",frame = "none")
nodelabels(round(convar3[ind_large3],3),large_node3,col="red",frame = "none")
```
Since alpha is unknown, the variance of contrasts being estimated would be different with varied alpha values. So, I split contrast tables into two different sets of contrasts based on either alpha has been underestimated or overestimated. 
```{r}
#explore alpha
n_sim=100000
mean(alpha_table3)
hist(alpha_table3,xlim=c(0,5))
alpha_underest=rep(2,n_sim)
alpha_underest[alpha_table3<eModel$alpha]=1
contrast_underest=subset(contrast_table3,alpha_underest==1)
contrast_overest=subset(contrast_table3,alpha_underest==2)
contrast_underest_var=apply(contrast_underest,2,var)
contrast_overest_var=apply(contrast_overest,2,var)
hist(contrast_underest_var)
hist(contrast_overest_var)
#Try to find the contrast with variance relatively small and large with underestimated alpha
ind_small_under=which(contrast_underest_var<0.04)
ind_median_under=which(contrast_underest_var<0.06&contrast_underest_var>=0.04)
ind_large_under=which(contrast_underest_var>=0.08)
median_node3_under=intnode[ind_median_under]
small_node3_under=intnode[ind_small_under]
large_node3_under=intnode[ind_large_under]
large_node3_under
```

Try to find the contrast`s variance with underestimated alpha
``````{r myplot3,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
par(oma=c(0,0,1.5,0))
plot(eModel$tree,show.tip.label = F  ,font=8,no.margin=T)
mtext(expression(paste(hat(a),"<", a)), outer=TRUE)
nodelabels(round(contrast_underest_var[ind_small_under],3),small_node3_under,col="blue",frame = "none",cex=0.6)
edgelabels(edge=shift_config,col="green",frame = "none",pch=17)
nodelabels(round(contrast_underest_var[ind_median_under],3),median_node3_under,col="purple",frame = "none",cex=0.6)
#nodelabels(round(contrast_underest_var[ind_large_under],3),large_node3_under,col="red",frame = "none")

```
Many contrasts with at least one child node as tip of the tree are underestimated. As mentioned above, contrasts of nodes with a shift edge are underestimated. One of the node labled in red has relatively large variance of contrast. This situation cannot be fully explained by its location closer to the root. (need some investigations)  


Try to find the contrast`s variance with overestimated alpha
```{r}
ind_small_over=which(contrast_overest_var<0.04)
ind_median_over=which(contrast_overest_var<0.06&contrast_overest_var>=0.04)
ind_large_over=which(contrast_overest_var>=0.08)
median_node3_over=intnode[ind_median_over]
small_node3_over=intnode[ind_small_over]
large_node3_over=intnode[ind_large_over]
large_node3_over
```





``````{r myplot4,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
par(oma=c(0,0,1.5,0))
plot(eModel$tree,show.tip.label = F  ,font=8,no.margin=T,direction = "leftwards")
mtext(expression(paste(hat(a),">", a)), outer=TRUE)
nodelabels(round(contrast_overest_var[ind_small_over],3),small_node3_over,col="blue",frame = "none",cex=0.6)
edgelabels(edge=shift_config,col="green",frame = "none",pch=17)
nodelabels(round(contrast_overest_var[ind_median_over],3),median_node3_over,col="purple",frame = "none",cex=0.6)
nodelabels(round(contrast_overest_var[ind_large_over],3),large_node3_over,col="red",frame = "none",cex=0.6)

```
When alpha has been overestimated, there are much more nodes with large variance of contrasts. And most of those nodes are closer to root. 


### Results from unknown shift configuration
	
and also unknown alpha and unknown beta. Spent 24hrs, generated 1262 eModels
```{r myplot5,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
load("unknownconfig_alpha_4.1.RData")
load("unknownconfig_shifts_4.1.RData")
load("unknownconfig_sigma2_4.1.RData")
load("unknownconfig_mu_4.1.RData")
load("unknownconfig_shiftvalues_4.1.RData")
load("unknownconfig_contrast_4.1.RData")
#Calculate mean of contrasts
contrast_mean4=colMeans(contrast_table4)
head(contrast_mean4) 
max(abs(contrast_mean4))
hist(contrast_mean4)
#Calculate variance of contrasts
convar4=apply(contrast_table4,2,ss)
hist(convar4)

#Shifts that occur the most often
vectorOfShift_1=unlist(vectorOfShift, recursive = TRUE, use.names = F)
freq_table=as.data.frame(table(vectorOfShift_1))
freq_table
```

Try to find the contrast with variance relatively small, median and large. 

```{r}
ind_small4=which(convar4<0.04)
ind_median4=which(convar4<0.06&convar4>=0.04)
ind_large4=which(convar4>=0.085)
median_node4=intnode[ind_median4]
small_node4=intnode[ind_small4]
large_node4=intnode[ind_large4]
large_node4

mean(convar4[-as.numeric(large_node4)])
```
Even though we ignore those nodes with large variance of contrasts, the variances of contrasts on average are still overestimated.  
```{r myplot6,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
mean_large4=intnode[which(abs(contrast_mean4)>0.05)]
median_node4=intnode[ind_median4]
small_node4=intnode[ind_small4]
large_node4=intnode[ind_large4]
plot(eModel$tree,show.tip.label = F,main="",font=8)
nodelabels(round(convar4[ind_small4],3),small_node4,col="blue",frame = "none")
for(i in 1:length(shift_config)){
edgelabels(paste(shift_config[i],"  ",round(freq_table$Freq[which(freq_table$vectorOfShift_1==shift_config[i])]/8,3),"%"),shift_config[i],col="green",frame = "none")
}
#56,73

edgelabels(paste(round(freq_table$Freq[which(freq_table$vectorOfShift_1==56)]/8,3),"%"),56,col="green",frame="none")
edgelabels(paste(round(freq_table$Freq[which(freq_table$vectorOfShift_1==73)]/8,3),"%"),73,col="green",frame = "none")
nodelabels(round(convar4[ind_median4],3),median_node4,col="purple",frame = "none")
nodelabels(round(convar4[ind_large4],3),large_node4,col="red",frame = "none",adj=c(0.5,0))
nodelabels(round(contrast_mean4[which(abs(contrast_mean4)>0.05)],3),mean_large4,col="black",frame = "none",adj=c(0.5,1))
mtext("blue: contrast variance < 0.04, purple: contrast variance > 0.04 and < 0.06,
      green dots: shifts",line=-21,side=3,las=0)
```
The shifts being detected most often are: 32,77,118,164,98,14,55,74 which are exactly corresponding to the shift configuration `r eModel$shift.configuration`of the true model. 
A few nodes with shift edges have larger than average mean of contrasts. So, their variances are overestimated accordingly. 

Contrasts of node with shift edges being detected or not
```{r}

contrast_plot = function(edge, contrast){
  contrast_detect=rep(NA, 600)
  for(i in 1:length(vectorOfShift)){
    if(edge%in%vectorOfShift[[i]]){
      contrast_detect[i]=TRUE
    } 
    else {
      contrast_detect[i]=FALSE
    }
  }
  plot(contrast_table4[,contrast] ~ factor(contrast_detect), xlab="detected?",ylab=paste(contrast,"th contrast value"))
  mtext(paste("edge: ",edge," contrast: ",contrast,
              ", detected: ", round(100*sum(contrast_detect)/800, 1), "%", sep=""))
  abline(h=0, lty=3)
  convar_bygroup = tapply(contrast_table4[,contrast], contrast_detect, var)
  mtext(side=1, at=1:2, text=round(convar_bygroup,4), line=-2)
}
```
Those figures show below represent contrasts of different nodes 
with group1(shift detected) and group2(shift not detected). 
```{r}
layout(matrix(1:4,2,2))
contrast_plot(14,7) # edge 14, node 193, contrast 7
contrast_plot(55,28)# edge 55, node 172, contrast 28
contrast_plot(74,37) # edge 74, node 163, contrast 37
contrast_plot(77,39) # edge 77, node 161, contrast 39

contrast_plot(98,49) # edge 98, node 151, contrast 49
contrast_plot(118,59)# edge 118, node 141, contrast 59
contrast_plot(164,82)# edge 164, node 118, contrast 82
contrast_plot(32,16)# last edge 32, node 184 , contrast 16

```
It is clear to see nodes with shift edges being detected have mean of contrasts close to zero. In some cases when those shift edges have not been detected have very different contrast values. Most of contrasts with shift edge not being detected have larger variances than variance of contrasts with shift edges being detected.  Edge 77 is especially strange because it does not follow last rule. (Need some explaination)





```{r myplot7,echo=TRUE, fig.width=8, fig.height=12, dpi=300}
load("unknownconfig_alpha_4.2.RData")
load("unknownconfig_shifts_4.2.RData")
load("unknownconfig_sigma2_4.2.RData")
load("unknownconfig_mu_4.2.RData")
load("unknownconfig_shiftvalues_4.2.RData")
load("unknownconfig_contrast_4.2.RData")
#Calculate mean of contrasts
contrast_mean5=colMeans(contrast_table5)
head(contrast_mean5) 
max(abs(contrast_mean5))
hist(contrast_mean5)
#Calculate variance of contrasts
convar5=apply(contrast_table5,2,ss)
hist(convar5)

#Shifts that occur the most often
vectorOfShift_2=unlist(vectorOfShift2, recursive = TRUE, use.names = F)
freq_table2=as.data.frame(table(vectorOfShift_2))
freq_table2
```

Try to find the contrast with variance relatively small, median and large. 

```{r}
ind_small5=which(convar5<0.04)
ind_median5=which(convar5<0.06&convar5>=0.04)
ind_large5=which(convar5>=0.085)
median_node5=intnode[ind_median5]
small_node5=intnode[ind_small5]
large_node5=as.numeric(intnode[ind_large5])
large_node5

mean(convar5[-large_node5])
```
Even though we ignore those nodes with large variance of contrasts, the variances of contrasts on average are still overestimated.  
```{r}
mean_large5=intnode[which(abs(contrast_mean5)>0.05)]
median_node5=intnode[ind_median5]
small_node5=intnode[ind_small5]
large_node5=intnode[ind_large5]
plot(eModel$tree,show.tip.label = F,main="small,median,large contrast nodes",font=8)
nodelabels(round(convar5[ind_small5],3),small_node5,col="blue",frame = "none")
for(i in 1:length(shift_config)){
edgelabels(paste(shift_config[i],"  ",round(freq_table2$Freq[which(freq_table2$vectorOfShift_1==shift_config[i])]/6,3),"%"),shift_config[i],col="green",frame = "none")
}
```
#56,73
```{r myplot8, eval=FALSE, fig.height=12, fig.width=8, dpi=300, include=FALSE}
edgelabels(paste(round(freq_table2$Freq[which(freq_table2$vectorOfShift_1==56)]/6,3),"%"),56,col="green",frame="none")
edgelabels(paste(round(freq_table2$Freq[which(freq_table2$vectorOfShift_1==73)]/6,3),"%"),73,col="green",frame = "none")
nodelabels(round(convar5[ind_median5],3),median_node5,col="pink",frame = "none")
nodelabels(round(convar5[ind_large5],3),large_node5,col="red",frame = "none",adj=c(0.5,0))
nodelabels(round(contrast_mean5[which(abs(contrast_mean5)>0.05)],3),mean_large5,col="black",frame = "none",adj=c(0.5,1))
```
The shifts being detected most often are: 32,77,118,164,98,14,55,74 which are exactly corresponding to the shift configuration `r eModel$shift.configuration`of the true model. 
A few nodes with shift edges have larger than average mean of contrasts. So, their variances are overestimated accordingly. 

Contrasts of node with shift edges being detected or not



```{r,eval=FALSE}

contrast_plot = function(edge, contrast){
  contrast_detect=rep(NA, 600)
  for(i in 1:length(vectorOfShift2)){
    if(edge%in%vectorOfShift2[[i]]){
      contrast_detect[i]=TRUE
    } 
    else {
      contrast_detect[i]=FALSE
    }
  }
  plot(contrast_table5[,contrast] ~ factor(contrast_detect), xlab="detected?",ylab=paste(contrast,"th contrast value"))
  mtext(paste("edge: ",edge," contrast: ",contrast,
              ", detected: ", round(100*sum(contrast_detect)/800, 1), "%", sep=""))
  abline(h=0, lty=3)
  convar_bygroup = tapply(contrast_table5[,contrast], contrast_detect, var)
  mtext(side=1, at=1:2, text=round(convar_bygroup,4), line=-2)
}
```
Those figures show below represent contrasts of different nodes 
with group1(shift detected) and group2(shift not detected). 


```{r,eval=FALSE}
layout(matrix(1:4,2,2))
contrast_plot(14,7) # edge 14, node 193, contrast 7
contrast_plot(55,28)# edge 55, node 172, contrast 28
contrast_plot(74,37) # edge 74, node 163, contrast 37
contrast_plot(77,39) # edge 77, node 161, contrast 39

contrast_plot(98,49) # edge 98, node 151, contrast 49
contrast_plot(118,59)# edge 118, node 141, contrast 59
contrast_plot(164,82)# edge 164, node 118, contrast 82
contrast_plot(32,16)# last edge 32, node 184 , contrast 16

```


```{r}
##layout for 2
layout(matrix(c(1,2,3,2), 2, 2, byrow = TRUE), respect = TRUE)
##layout for 3,4
layout(matrix(1:2))

 

```