---
title: "Estimate contrast with known alpha"
author: "Qing (Sabrina) Yu"
date: "Feb 22,2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal of the study

In this report, our goal is to check the distribution of phylogenetic contrasts
with known alpha, known shift configuration and unknown shift values that are used in the bootstrap procedure 
in the `l1ou` package. In the last report, we used the true shift edge, true shift values, true alpha.

We used code from the bootstrap procedure to simulate contrast values.
Then we calculated the mean and standard deviation of the contrast values at each node.
The conclusions above held for all nodes, whether nodes had a shift on a child edge or not. 


#  Estimate shift configuration from orignal lizard data (change)

We used the lizard tree from the `phylolm` package and associated trait data on these lizard
species (just the first trait, which is the first PC axis from a PCA).


```{r,include=FALSE}
library(l1ou)
library(phylolm)
data(lizard.tree, lizard.traits)
lizard <- adjust_data(lizard.tree, lizard.traits[,1])
n_tips=length(lizard$tree$tip.label) # Total number of tips
eModel <- estimate_shift_configuration(lizard$tree, lizard$Y)
save(eModel,  file = "eModel_2_27.RData")
```

This trait was analyzed to estimate the shifts in trait evolution using the function
`fit_OU` from the `l1ou` package. 

```{r}
load("eModel_2_27.RData")
data(lizard.tree, lizard.traits)
lizard <- adjust_data(lizard.tree, lizard.traits[,1])
truealpha=eModel$alpha
y0=eModel$intercept
truetheta = y0 + l1ou:::convert_shifts2regions(eModel$tree, 
                                eModel$shift.configuration, eModel$shift.values)
nShifts   = length(eModel$shift.configuration) # Total number of shifts
n_tips=length(eModel$tree$tip.label) # Total number of tips
sigma2=eModel$sigma2
shiftnode= eModel$tree$edge[eModel$shift.configuration,1]-n_tips # internal nodes with shift
othernode=(1:(n_tips-1))[-shiftnode] # Other nodes: those without a true shift
```


We used this model to simulate new data using the `rTraitCont` function from the `phylolm` package.
Below, `RE` is the result of the function `sqrt_OU_covariance`, which calculates the square-root of the 
phylogenetic covariance matrix with a recursive algorithm, which traverses the tree once.
`covInverseSqrt` is the inverse square root of the phylogenetic covariance matrix, and `covSqrt` is the square-root of the phylogenetic covariance matrix. 
Finally, `contrast` contains the contrasts at all nodes. 
These matrices and contrasts were obtained using the true value of $\alpha$,
the same value used to simulate the data. This is an ideal situation when
$\alpha$ is known without error.

# simulation procedure

we use the same data for all scenarios:
set.seed here, then save data in ytable.
```{r,eval=F}
n_sim=100000
Y_table=matrix(nrow=n_sim, ncol=n_tips, data=NA)
set.seed(1293)
for (i in 1:n_sim) {
  Y  <- rTraitCont(eModel$tree, "OU", theta=truetheta, 
                 alpha=truealpha, 
                 sigma=sqrt(eModel$sigma2), root.value=y0)
  Y_table[i,]=Y
}
save(Y_table, file="Y_table_2_27.RData")
```

```{r}
load("Y_table_2_27.RData")
len=dim(Y_table)[1]
```
# Estimation Procedure

We consider 4 scenarios for the calculation of contrasts, depending
on what parameters are known (yes) or unknown (no) for the calculation
of the inverse square-root of the covariance matrix.

parameter |  1  |  2 (knownalpha)  |  3 () |  4 ()
----------|-----|-----|-----|-----
β<sub>0</sub>, β<sub>1</sub>, ...| yes | no | no | no
α | yes | yes | no | no
shift config | yes | yes | yes | no

For the tree topology, we assume it known always.
Refer to "filename" for the results of simulation 1, in which we know everything,
and the contrasts behave as expected.

# scenario 1: all parameters known and set to their true values

use Y table, call contrasts table "contrast_table1"
```{r,eval=F}
REf = sqrt_OU_covariance(lizard.tree, alpha=truealpha, 
                        root.model =  eModel$l1ou.options$root.model,normalize.tree.height=T ,
                        check.order=F, check.ultrametric=F)

Dtf  = t(REf$sqrtInvSigma)
contrast_table=matrix(nrow=n_sim, ncol=n_tips, data=NA)
for (i in 1:nrow(Y_table)) {
  contrast_table[i,]=Dtf%*%(Y_table[i,] - eModel$mu)
}
save(contrast_table, file="contrast_table_2_27.RData")
```



# scenario 2: known alpha and known shift configuration, beta unknown around 3 hrs

```{r,eval=F,include=T}
RE = sqrt_OU_covariance(lizard$tree, alpha=eModel$alpha, # assumed known here
                        root.model = "OUfixedRoot",
                        check.order=F, check.ultrametric=F)
covInverseSqrt  <- t(RE$sqrtInvSigma)

contrast_table2=matrix(nrow=len, ncol=n_tips, data=NA)
sigma_table2=rep(NA,len)
mu_table2=matrix(nrow=len, ncol=n_tips, data=NA)
shift_values2=matrix(nrow=len, ncol=length(shift_config), data=NA)
for (i in 1:len) {
  model= fit_OU(lizard$tree, Y_table[i,], shift_config,
                alpha.upper=truealpha, alpha.lower=truealpha, alpha.starting.value=truealpha)
  # the line above is to fix alpha to the known value
 sigma_table2[i] = model$sigma2
 shift_values2[i,]=model$shift.values
 mu_table2[i,]=model$mu
  contrast_table2[i,]  <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(sigma_table2,file="knownalpha_sigma2.RData")
 save(contrast_table2,file="knownalpha_contrast.RData")
 save(mu_table2,file="knownalpha_mu.RData")
 save(shift_values2,file="knownalpha_shiftvalues.RData")
```

# scenario 3: known shiftconfig but unknown alpha and unknown beta

```{r,eval=F,include=T}
contrast_table3 = matrix(nrow=len, ncol=n_tips, data=NA)
alpha_table3  = rep(NA,len)
sigma_table3 = rep(NA,len)
mu_table3 = matrix(nrow=len, ncol=n_tips, data=NA)
shift_values3 = matrix(nrow=len, ncol=length(shift_config), data=NA)
# also save sigma2 estimated values, and 8 shift values? and/or 100 mu values?
for (i in 1:len) {
  model= fit_OU(lizard$tree, Y_table[i,], shift_config)
  alpha_table3[i] = model$alpha
  sigma_table3[i] = model$sigma2
  shift_values3[i,]=model$shift.values
  mu_table3[i,]=model$mu
  RE = sqrt_OU_covariance(lizard$tree, alpha=model$alpha, # alpha estimated here
                          root.model = "OUfixedRoot",
                          check.order=F, check.ultrametric=F)
  covInverseSqrt  <- t(RE$sqrtInvSigma)
  contrast_table3[i,] <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(contrast_table3,file="unknownalpha_contrast.RData")
save(alpha_table3, file="unknownalpha_alpha.RData")
save(sigma_table3,file="unknownalpha_sigma2.RData")
save(mu_table3,file="unknownalpha_mu.RData")
save(shift_values3,file="unknownalpha_shiftvalues.RData")
```

# scenario 4: unknown shift configuration, unknown alpha and unknown beta

```{r,eval=F,include=T}
n_sim=2000 # lower because estimation is a lot slower when we have to search for the config
# use first n_sim rows of Ytable only
Y_table=Y_table[1:n_sim,]
contrast_table4=matrix(nrow=n_sim, ncol=n_tips, data=NA)
vectorOfShift <- vector(mode = "list", length = n_sim)
alpha_table4  = rep(NA,n_sim)
sigma_table4 = rep(NA,n_sim)
mu_table4 = matrix(nrow=n_sim, ncol=n_tips, data=NA)
shift_values4 = vector(mode = "list", length = n_sim)

for (i in 1:n_sim) {
  model= estimate_shift_configuration(lizard$tree, Y_table[i,])
  vectorOfShift[[i]] =model$shift.configuration
  alpha_table4[i] = model$alpha
  sigma_table4[i] = model$sigma2
  shift_values4[[i]]=model$shift.values
  mu_table4[i,]=model$mu
  RE = sqrt_OU_covariance(lizard$tree ,alpha=model$alpha,
                          root.model = "OUfixedRoot",
                          check.order=FALSE, check.ultrametric=FALSE)
  covInverseSqrt  <- t(RE$sqrtInvSigma)
  contrast_table4[i,] <-  covInverseSqrt%*%(Y_table[i,] - model$mu)
}
save(contrast_table4,file="unknownconfig_contrast.RData")
save(alpha_table4, file="unknownconfig_alpha.RData")
save(sigma_table4,file="unknownconfig_sigma2.RData")
save(mu_table4,file="unknownconfig_mu.RData")
save(shift_values4,file="unknownconfig_shiftvalues.RData")
save(vectorOfShift,file="unknownconfig_shifts.RData")
```

# Results: visualizations of contrast distributions

Function to calculate the variance of contrasts assuming mean 0.
```{r}
ss=function(x){
  sum(x^2)/length(x)
}
```
###all parameters known and set to their true values
```{r}
 load("~/R/traitevoOUshift/contrast_cases/allknown_contrast.RData")
head(colMeans(contrast_table1)) 
max(abs(colMeans(contrast_table1)))
round(colMeans(contrast_table1),5)
convar=apply(contrast_table1,2,ss)
round(convar,3)
mean(convar)
sigma2
```
Mean of contrasts are close to zero but the variance of contrasts is not equal to `sigma2`.
###known alpha and known shift configuration, beta unknown
```{r}
load("~/R/traitevoOUshift/contrast_cases/knownalpha_contrast.RData")
load("~/R/traitevoOUshift/contrast_cases/knownalpha_sigma2.RData")
load("~/R/traitevoOUshift/contrast_cases/knownalpha_shiftvalues.RData")
load("~/R/traitevoOUshift/contrast_cases/knownalpha_mu.RData")
head(colMeans(contrast_table2)) 
max(abs(colMeans(contrast_table2)))
round(colMeans(contrast_table2),5)
convar=apply(contrast_table2,2,ss)
hist(convar)
round(convar,3)
mean(convar)
sigma2
#Check which contrasts are extremely different from 0
```
###known shiftconfig but unknown alpha and unknown beta
```{r}
load("~/R/traitevoOUshift/contrast_cases/unknownalpha_contrast.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownalpha_sigma2.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownalpha_shiftvalues.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownalpha_mu.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownalpha_alpha.RData")
head(colMeans(contrast_table3)) 
max(abs(colMeans(contrast_table3)))
round(colMeans(contrast_table3),5)
convar=apply(contrast_table3,2,ss)
round(convar,3)
mean(convar)
sigma2
#Check which contrasts are extremely different from 0
```
###unknown shift configuration, unknown alpha and unknown beta spent 24hrs, generated 1262 eModels
```{r}
load("~/R/traitevoOUshift/contrast_cases/unknownconfig_alpha.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownconfig_shifts.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownconfig_sigma2.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownconfig_mu.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownconfig_shiftvalues.RData")
load("~/R/traitevoOUshift/contrast_cases/unknownconfig_contrast.RData")
#mean of contrasts are extremely different from zero
head(colMeans(contrast_table4)) 
max(abs(colMeans(contrast_table4)))
colMeans(contrast_table4)
convar=apply(contrast_table4,2,ss)
convar
mean(convar)
head(vectorOfShift)
#Shifts that occur the most often
vectorOfShift=unlist(vectorOfShift, recursive = TRUE, use.names = F)
as.data.frame(table(vectorOfShift))

```
The shifts being detected most often are: 32,77,118,164,98,14,55,74 which are exactly corresponding to the shift configuration `r shift_config`of the true model. 